Model-based Reinforcement Learning with Fast Posterior Updates

### 2023-03-09 paper/main.tex
commit after adding some SVGO maths and algs

% Our packages
\usepackage{todonotes}
\usepackage[colorlinks=true,linkcolor=blue,allcolors=blue]{hyperref}
\usepackage{amsmath}
\usepackage{bm}

@@

@@

\section{Problem Statement and Background} \label{sec:problem-statement}
We consider environments with states \(s \in \mathcal{S}\), actions \(a \in \mathcal{A}\) and transition dynamics \(f: \mathcal{S} \times \mathcal{A} \rightarrow \mathcal{S}\) subject to
iid noise \(\epsilon_{t}\), given by,
\begin{align}
s_{t+1} = f(s_{t}, a_{t}) + \epsilon_{t}.
\end{align}

\textbf{Reinforcement learning RL}
The goal of reinforcement learning is to find a policy \(\pi \in \Pi\) that maximises the sum of discounted
rewards in expecation under the transition noise (aleatoric uncertainty),
\begin{align} \label{eq-model-free-objective}
\pi^{*} = \arg \max_{\pi \in \Pi} J(f, \pi) = \arg \max_{\pi \in \Pi} \mathbb{E}_{\bm\epsilon_{0:\infty}} \left[ \sum_{t=0}^{\infty} \gamma^{t} r(s_{t},a_{t}) \right].
\end{align}

@@

\section{Experiments}
Environments
\begin{itemize}
  \item Compare updated posterior with non updated posterior
  \begin{itemize}
  \item Compare for different strategies of making decisions under dynamic model's uncertainty:
    \begin{itemize}
      \item \textbf{Greedy exploitation} \(\pi_{\text{greedy}}\)
      \item \textbf{Hallucinated-UCRL} \(\pi_{\text{HUCRL}}\)
      \item \textbf{Thompson sampling} \(\pi_{\text{TS}}\)
    \end{itemize}
  \end{itemize}
  \item Compare impact of stationary vs non-stationary priors
  \begin{itemize}
    \item First do this with GP
    \item Then try to do with Laplace BNN
  \end{itemize}
\end{itemize}

>>>>

% Our packages
\usepackage{todonotes}
\usepackage[colorlinks=true,linkcolor=blue,allcolors=blue]{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{algpseudocode}
\usepackage{algorithm}

@@

% Variables
\newcommand{\state}{\ensuremath{\mathbf{s}}}
\newcommand{\action}{\ensuremath{\mathbf{a}}}
\newcommand{\noise}{\ensuremath{\bm\epsilon}}
\newcommand{\discount}{\ensuremath{\gamma}}
\newcommand{\inducingVariable}{\ensuremath{\mathbf{u}}}
\newcommand{\dataset}{\ensuremath{\mathcal{D}}}
\newcommand{\dualParam}[1]{\ensuremath{\bm{\lambda}_{#1}}}
\newcommand{\meanParam}[1]{\ensuremath{\bm{\mu}_{#1}}}

% Indexes
\newcommand{\horizon}{\ensuremath{h}}
\newcommand{\Horizon}{\ensuremath{H}}

% Domains
\newcommand{\stateDomain}{\ensuremath{\mathcal{S}}}
\newcommand{\actionDomain}{\ensuremath{\mathcal{A}}}
\newcommand{\policyDomain}{\ensuremath{\Pi}}

% Functions
\newcommand{\rewardFn}{\ensuremath{r}}
\newcommand{\transitionFn}{\ensuremath{f}}
\newcommand{\latentFn}{\ensuremath{f}}

\newcommand{\optimisticTransition}{\ensuremath{\hat{f}}}
\newcommand{\optimisticTransitionMean}{\ensuremath{\mu_{\optimisticTransition}}}
\newcommand{\optimisticTransitionCov}{\ensuremath{\mu_{\optimisticTransition}}}
\newcommand{\optimisticTransitionSet}{\ensuremath{\mathcal{M}}}

% Parameters
\newcommand{\transitionParams}{\ensuremath{\phi}}
\newcommand{\valueFnParams}{\ensuremath{\psi}}
\newcommand{\policyParams}{\ensuremath{\theta}}

% Networks
\newcommand{\transition}{\ensuremath{\mathbf{d}_{\transitionParams}}}
\newcommand{\valueFn}{\ensuremath{\mathbf{Q}}}
\newcommand{\stateValueFn}{\ensuremath{\mathbf{V}}}
% \newcommand{\valueFn}{\ensuremath{\mathbf{Q}_{\valueFnParams}}}
\newcommand{\policy}{\ensuremath{\pi}}
\newcommand{\pPolicy}{\ensuremath{\pi_{\policyParams}}}

@@

\section{Problem Statement and Background} \label{sec:problem-statement}
We consider environments with states \(\state \in \stateDomain \), actions \(\action \in \actionDomain\) and transition dynamics \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \) subject to
iid noise \(\noise_{t}\), given by,
\begin{align}
\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}.
\end{align}

\textbf{Reinforcement learning (RL)}
The goal of reinforcement learning is to find a policy \(\pi \in \Pi\) that maximises the sum of discounted
rewards in expecation under the transition noise (aleatoric uncertainty),
\begin{align} \label{eq-model-free-objective}
\policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \left[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \right].
\end{align}

@@

\begin{align} \label{eq-fast-update-mpc}
  \policy(\state) = \arg &\max_{\action_{0}} \max_{\action_{1}, \ldots, \action_{\Horizon}}
  \mathbb{E}_{\state_{\horizon} \sim p(\state_{\horizon+1} \mid \transitionFn(\state_{\horizon}, \action_{\horizon}))} \left[ \sum_{\horizon=0}^{\Horizon}     \discount^{\horizon} \rewardFn(\state_{\horizon},\action_{\horizon}) \mid \state_{0}=\state \right] + \discount^{\Horizon+1} \stateValueFn(\state_{\Horizon+1})
\end{align}

\begin{align} \label{eq-fast-update-mpc}
  \stateValueFn(\state) = \mathbb{E} \left[ \sum_{t=0}^{\infty}     \discount^{t} \rewardFn(\state_{t},\action_{t}) \mid \state_{0}=\state \right]
\end{align}

\begin{align} \label{eq-fast-update-mpc}
  \policy(\state) = \arg &\max_{\action_{0}} \max_{\action_{1}, \ldots, \action_{\Horizon}} \max_{\optimisticTransition \in \optimisticTransitionSet}
  \sum_{\horizon=0}^{\Horizon}  \mathbb{E}_{\noise_{\horizon}} \left[  \discount^{\horizon} \rewardFn(\state_{\horizon},\action_{\horizon}) \right] + \discount^{\Horizon+1} \stateValueFn(\state_{\Horizon+1}) \\
  \text{s.t. } \state_{\horizon+1} &= \optimisticTransition(\state_{\horizon}, \action_{\horizon}) + \noise_{\horizon} \\
  \optimisticTransition(\state_{\horizon}, \action_{\horizon}) &=
\optimisticTransitionMean(\state_{\horizon}, \action_{\horizon}) \pm \beta_{i}
\optimisticTransitionCov(\state_{\horizon}, \action_{\horizon})
\end{align}

\begin{align} \label{eq-svgp-predictive-posterior}
  q_{\inducingVariable}(\transitionFn(\state_{t}, \action_{t})) = \mathcal{N}
  \left( \transitionFn(\state_{t}, \action_{t}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} \mathbf{A}^{T} + \mathbf{A} \mathbf{V}^{*} \mathbf{A}^{T} \right)
\end{align}

\begin{align} \label{eq-matrix-A}
  \mathbf{A} = \mathbf{K}_{\mathbf{X}\mathbf{Z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}}
\end{align}

\begin{align} \label{eq-dual-params}
  \mathbf{m}^{*} = \mathbf{V}^{*}\dualParam{1}^{*} \quad \mathbf{V}^{*} = [\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} + \dualParam{2}^{*}]
\end{align}

\begin{align} \label{eq-dual-update-svgp-old-new}
 \dualParam{1}^{\text{new}} &\leftarrow \dualParam{1}^{\text{old}} +
  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\text{new}}, \action_{\text{new}}))}
 \left[ \log p(\state_{\text{new}} \mid \latentFn(\state_{\text{new}}, \action_{\text{new}}) ) \right] \\
 \dualParam{2}^{\text{new}} &\leftarrow \dualParam{2}^{\text{old}} +
  \nabla_{\meanParam{2}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\text{new}}, \action_{\text{new}}))}
 \left[ \log p(\state_{\horizon+1} \mid \latentFn(\state_{\text{new}}, \action_{\text{new}}) ) \right]
\end{align}

\begin{align} \label{eq-dual-update-svgp}
 \dualParam{1}^{\horizon+1} &\leftarrow \dualParam{1}^{\horizon} +
  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\horizon}, \action_{\horizon}))}
 \left[ \log p(\state_{\horizon+1} \mid \latentFn(\state_{\horizon}, \action_{\horizon}) ) \right] \\
 \dualParam{2}^{\horizon+1} &\leftarrow \dualParam{2}^{\horizon} +
  \nabla_{\meanParam{2}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\horizon}, \action_{\horizon}))}
 \left[ \log p(\state_{\horizon+1} \mid \latentFn(\state_{\horizon}, \action_{\horizon}) ) \right]
\end{align}

\begin{algorithm}[H]
\caption{Model-based RL}\label{alg-mbrl}
\begin{algorithmic}[1]
    % ${p(\state_{\timeInd+1} \mid \singleInput, \dataset_{0})}$}
\For{$i  \in \{0, 1, \ldots, \text{num episodes} \}$}
    \State Train dynamics using $\dataset_{0:i}$
    \For{$t  \in \{0, 1, \ldots, \text{num steps} \}$}
      \State Execute policy $\policy(\state_{t})$ in environment
      \State Set $\mathcal{D}_{i} = \mathcal{D}_{i} \cup (\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
    \EndFor
    \State Update data $\mathcal{D}_{0:i} = \mathcal{D}_{0:i-1} \cup \mathcal{D}_{i}$
\EndFor
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[H]
\caption{Model-based RL with fast updates}\label{alg-mbrl-fast-updates}
\begin{algorithmic}[1]
    % ${p(\state_{\timeInd+1} \mid \singleInput, \dataset_{0})}$}
\For{$i  \in \{0, 1, \ldots, \text{num episodes} \}$}
    \State Train dynamics using $\dataset_{0:i}$
    \For{$t  \in \{0, 1, \ldots, \text{num steps} \}$}
      \State Execute policy $\policy(\state_{t})$ in environment
      \State Set $\mathcal{D}_{i} = \mathcal{D}_{i} \cup (\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
      \State {\color{blue}Update dynamics posterior}
    \EndFor
    \State Update data $\mathcal{D}_{0:i} = \mathcal{D}_{0:i-1} \cup \mathcal{D}_{i}$
\EndFor
\end{algorithmic}
\end{algorithm}

\begin{minipage}{0.499\textwidth}
\begin{algorithm}[H]
\caption{Model-based RL}\label{alg-mbrl}
\begin{algorithmic}[1]
  % \Require{Start state $\state_0$, initial data set $\dataset_0$, policy $\pi_{0}$, dynamic model}
    % ${p(\state_{\timeInd+1} \mid \singleInput, \dataset_{0})}$}
\For{$i  \in \{0, 1, \ldots, \text{num episodes} \}$}
    \State Train dynamics using $\dataset_{0:i}$
    \State Train value function $\stateValueFn$ using $\dataset_{0:i}$
    \For{$t  \in \{0, 1, \ldots, \text{num steps} \}$}
      % \State Calculate $\action_{t}$ using \cref{eq-fast-update-mpc}
      % \State Execute 1 step in environment using $\policy(\state_{t})$ and observe $(\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
      \State Execute policy $\policy(\state_{t})$ in environment
      \State Set $\mathcal{D}_{i} = \mathcal{D}_{i} \cup (\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
    \EndFor
    \State Update data $\mathcal{D}_{0:i} = \mathcal{D}_{0:i-1} \cup \mathcal{D}_{i}$
\EndFor
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.499\textwidth}
\begin{algorithm}[H]
\caption{Model-based RL with fast updates}\label{alg-mbrl-with-fast-updates-simple}
\begin{algorithmic}[1]
  % \Require{Start state $\state_0$, initial data set $\dataset_0$, policy $\pi_{0}$, dynamic model}
    % ${p(\state_{\timeInd+1} \mid \singleInput, \dataset_{0})}$}
\For{$i  \in \{0, 1, \ldots, \text{num episodes} \}$}
    \State Train dynamics using $\dataset_{0:i}$
    \State Train value function $\stateValueFn$ using $\dataset_{0:i}$
    \For{$t  \in \{0, 1, \ldots, \text{num steps} \}$}
      \State Execute policy $\policy(\state_{t})$ in environment
      \State Set $\mathcal{D}_{i} = \mathcal{D}_{i} \cup (\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
      \State {\color{blue}Update dynamics posterior}
    \EndFor
    \State Update data $\mathcal{D}_{0:i} = \mathcal{D}_{0:i-1} \cup \mathcal{D}_{i}$
\EndFor
\end{algorithmic}
\end{algorithm}
\end{minipage}

\begin{algorithm}[!t]
\caption{Model-based RL with fast updates}\label{alg-mbrl-with-fast-updates}
\begin{algorithmic}[1]
  % \Require{Start state $\state_0$, initial data set $\dataset_0$, policy $\pi_{0}$, dynamic model}
    % ${p(\state_{\timeInd+1} \mid \singleInput, \dataset_{0})}$}
\For{$i  \in \{0, 1, \ldots, \text{num episodes} \}$}
    \State Train dynamics using $\dataset_{0:i}$ to get $p(\mathbf{\transitionFn} \mid \dataset_{0:i})$
    \State Train value function $\stateValueFn$ using $\dataset_{0:i}$
    % \While{not converged}
    %     \State Sample $N_{b}$ state transitions $\mathcal{B} \sim \dataset_{0:i}$
    %     \State Update dynamics using with $\mathcal{B}$
    % \EndWhile
    \For{$t  \in \{0, 1, \ldots, \text{num steps} \}$}
      \State Calculate $\action_{t}$ using \cref{eq-fast-update-mpc}
      \State Execute $\action_{t}$ in environment and observe $(\state_{t}, \action_{t}, \state_{t+1}, r_{t+1})$
      \State Update dynamics posterior (i.e. fast update via dual parameters)
      \begin{align}
      \dualParam{1}^{\horizon+1} &\leftarrow \dualParam{1}^{\horizon} +  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\horizon}, \action_{\horizon}))} \left[ \log p(\state_{\horizon+1} \mid \latentFn(\state_{\horizon}, \action_{\horizon}) ) \right] \\
      \dualParam{2}^{\horizon+1} &\leftarrow \dualParam{2}^{\horizon} +  \nabla_{\meanParam{2}} \mathbb{E}_{q_{\inducingVariable}(\latentFn(\state_{\horizon}, \action_{\horizon}))}  \left[ \log p(\state_{\horizon+1} \mid \latentFn(\state_{\horizon}, \action_{\horizon}) ) \right]
      \end{align}
    \EndFor
    % \State Calculate $\action_{\horizon}$ using \cref{eq-fast-update-mpc}
    % \State Execute $\action_{\horizon}$ in environment and observe $(\state_{\horizon}, \action_{\horizon}, \state_{\horizon+1}, r_{\horizon+1})$
\EndFor
\end{algorithmic}
\end{algorithm}

\section{Experiments}
Environments
\begin{itemize}
  \item Compare updated posterior with non updated posterior
  \begin{itemize}
  \item Compare for different strategies of making decisions under dynamic model's uncertainty:
    \begin{itemize}
      \item \textbf{Greedy exploitation} \(\pi_{\text{greedy}}\)
      \item \textbf{Hallucinated-UCRL} \(\pi_{\text{HUCRL}}\)
      \item \textbf{Thompson sampling} \(\pi_{\text{TS}}\)
    \end{itemize}
  \end{itemize}
  \item Compare impact of stationary vs non-stationary priors
  \begin{itemize}
    \item First do this with GP
    \item Then try to do with Laplace BNN
  \end{itemize}
  \item Compare impact of function vs weight space
\end{itemize}
